---
title: "Practice Worksheet: Gerrymandering"
format: typst
---

Recall the 3 quantitative approaches to measuring gerrymandering discussed in lecture:\
1. Mean - Median Score (Grofman, 1983)\
2. Efficiency Gap (Stephanopolous and McGee, 2015)\
3. Ensembles / Extreme Outliers\
Use the following table to practice applying the first two techniques.\

| District | D Votes | R Votes | Result  |
|:---------:|:-------:|:-------:|:--------|
| 1 | 62 | 38 | D wins |
| 2 | 55 | 45 | D wins |
| 3 | 44 | 56 | R wins |
| 4 | 47 | 53 | R wins |
| 5 | 46 | 54 | R wins |
| **Total** | **254** | **246** |          |

```{r}
#| message: false
#| echo: false
library(tidyverse)
votes <- tibble(District = 1:5,
                d_votes = c(62, 55, 44, 47, 46),
                r_votes = c(38, 45, 56, 53, 54),
                result = c(rep("D wins", 2), rep("R wins", 3)))
```

### Q1. Mean-Median Score

Calculate the mean-median score associated with this particular district breakdown of votes.

```{r}
# add code for Q1 here
v = votes$d_votes
r = votes$r_votes

v_mean = mean(v)
v_median = median(v)

r_mean = mean(r)
r_median = median(r)
  
print(paste("v_mean is: ", v_mean))
print(paste("v_median is: ", v_median))
print(paste("r_mean is: ", r_mean))
print(paste("v_median is: ", v_median))

```

### Q2. Efficiency Gap

Calculate the Efficiency Gap (EG) associated with this particular district breakdown of votes.

*Tip*: it may be helpful to write a function that takes a vector of votes across districts for party A and vector of votes for the party B and returns a vector of the number votes wasted by party A across districts.


```{r}
# add code for Q2 here
total = v + r
print(total)

wasted_votes <- function(party, oppo) {
    total = party + oppo
    winner = ifelse(party > oppo, "party", "opponent")
    wasted = ifelse(winner == "party", party - (total / 2 + 1), party)
    return(wasted)
}
wasted_d = wasted_votes(v,r)
wasted_r = wasted_votes(r,v)

res = sum(wasted_d - wasted_r) / sum(total)
print(paste("EG:", res))

```